---
title: "A Rule Learning Approach for Detecting Faults in Highly Configurable Software Systems from Uniform Random Samples"
author: "Ruben Heradio (rheradio@issi.uned.es)"
date: "Date: `r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

This report reproduces the experiments corresponding to the paper:

Ruben Heradio, David Fernandez-Amoros, Victoria Ruiz, and Manuel J. Cobo. *A Rule Learning Approach for Detecting Faults in Highly Configurable Software Systems from Uniform Random Samples*. Hawaii International Conference on System Sciences (HICSS), 2022. Hawaii, USA.

In it, the [JHipster dataset](https://github.com/xdevroey/jhipster-dataset) is analyzed with the following rule induction engines: AQ, CN2, LEM2, PART, and RIPPER

# Initialization

## Importing packages and installing them if needed

```{r package_import, results='hide', warning=FALSE}
REQUIRED_PACKAGES <- 
  c("tidyverse",
    "RoughSets",
    "RWeka",
    "gmodels",
    "caret",
    "psych",
    "xtable", 
    "kableExtra")
lapply(
  REQUIRED_PACKAGES,
  function(pkg) {
    print(pkg)
    if (system.file(package = pkg) == "") {
      install.packages(pkg,
                       repos = "http://cran.us.r-project.org"
      )
    }
    do.call("library", list(pkg))
  }
)
sessionInfo()
```
```{r}
sessionInfo()
```

## Constants

```{r global-constants-and-variables-initialization}
TEST_RESULTS_FILE <- "../dataset/jhipster_3.6.1_test_results.csv"
RULE_INDUCTION_ALGORITHMS <- c("AQ", "LEM2", "CN2", "JRip", "PART")
```

# Defining the function that runs and quantifies the performance of the rule learning systems

```{r function-that-tests-a-rule-algorithm}
check_rule_induction_algorithms <-
  function(sample_sizes,
           repetitions = 1,
           rule_induction_algorithms = RULE_INDUCTION_ALGORITHMS,
           test_results_file = TEST_RESULTS_FILE) {
    test_results <- read.csv(test_results_file,
                             header = TRUE,
                             sep = ",")
    
    test_results <- test_results %>%
      map( ~ as_factor(.x)) %>%
      as_tibble()
    
    rule_results <-
      tibble(
        sample_size = numeric(),
        rule_algorithm = character(),
        Accuracy = numeric(),
        Sensitivity = numeric(),
        Specificity = numeric(),
        Precision = numeric(),
        Kappa = numeric(),
      )
    
    test_results_dt <-
      SF.asDecisionTable(
        test_results,
        decision.attr = ncol(test_results),
        indx.nominal = 1:ncol(test_results)
      )
    
    for (s in sample_sizes) {
      writeLines(str_c("Sample size = ", s))
      for (r in seq_len(repetitions)) {
        writeLines(str_c("   Repetition = ", r))
        sample_indices <- sample(nrow(test_results),
                                 size = s,
                                 replace = TRUE)
        sample <- test_results[sample_indices, ]
        for (ra in rule_induction_algorithms) {
          writeLines(str_c("Rule induct. algorithm = ", ra))
          
          if (ra %in% c("AQ", "LEM2", "CN2")) {
            sample_dt <-
              SF.asDecisionTable(
                sample,
                decision.attr = ncol(sample),
                indx.nominal = 1:ncol(sample)
              )
            rules <-
              eval(parse(text = str_c(
                "RI.", ra, "Rules.RST(sample_dt)"
              )))
            predicted <- predict(rules, test_results_dt)
            predicted$predictions <-
              as.factor(predicted$predictions)
            cm <-
              confusionMatrix(predicted$predictions,
                              test_results$Test,
                              positive = "KO")
            rule_results <- rule_results %>%
              add_row(
                sample_size = s,
                rule_algorithm = ra,
                Accuracy = cm$overall[[1]],
                Sensitivity = sensitivity(
                  predicted$predictions,
                  test_results$Test,
                  positive =
                    "KO"
                ),
                Specificity = specificity(
                  predicted$predictions,
                  test_results$Test,
                  positive =
                    "OK"
                ) ,
                Precision = posPredValue(
                  predicted$predictions,
                  test_results$Test,
                  positive =
                    "KO"
                ) ,
                Kappa = cm$overall[[2]],
              )
          } else if (ra %in% c("JRip", "PART")) {
            rules <-
              eval(parse(text = str_c(ra, "(Test ~., data=sample)")))
            predicted <- predict(rules, test_results)
            cm <-
              confusionMatrix(predicted, test_results$Test, positive = "KO")
            rule_results <- rule_results %>%
              add_row(
                sample_size = s,
                rule_algorithm = ra,
                Accuracy = cm$overall[[1]],
                Sensitivity = sensitivity(predicted, test_results$Test, positive =
                                            "KO"),
                Specificity = specificity(predicted, test_results$Test, positive =
                                            "OK") ,
                Precision = posPredValue(predicted, test_results$Test, positive =
                                           "KO") ,
                Kappa = cm$overall[[2]],
              )
          } else {
            stop(str_c("The rule induction algorithm ", ra, " is not supported"))
          }
        }
      }
    }
    rule_results
  }
```

# Run the experiments
```{r run-experiments}
rule_results <- check_rule_induction_algorithms(
                  sample_sizes=20:200
                )
```

# Experimental results

## Linear models that account for the influence of the rule induction algorithms and the sample size on the performance metrics

```{r linear-models}
for (m in c("Accuracy", "Sensitivity", "Specificity", "Precision", "Kappa")) {
  writeLines(m)
  eval(parse(
    text = str_c("with(rule_results, describeBy(", m, ", rule_algorithm))")
  ))
  eval(parse(
    text = str_c(
      "model <- lm(",
      m,
      " ~ -1+ rule_algorithm + sample_size, data=rule_results)"
    )
  ))
  print(summary(model))
}
```

## Plot showing how the performance measures depend on the sample size and the used rule induction algorithm
```{r plot-performance-measures-dependecy, warning=FALSE}
tidied_rule_results <- rule_results %>%
  gather(Accuracy:Kappa, key = Measure, value = Value) %>%
  mutate(Measure = fct_recode(Measure,
                              "Recall/Sensitivity" = "Sensitivity"))

ggplot(tidied_rule_results, aes(x = sample_size, y = Value)) +
  geom_point(alpha = 0.3, color = "#3388CC") +
  geom_smooth(se = FALSE, color = "#E91E63", size = 1) +
  facet_grid(Measure ~ rule_algorithm) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 200, by = 25)) +
  coord_cartesian(xlim = c(0, 200), ylim = c(0, 1)) +
  labs(x = expression("Sample size"),
       y = expression("Performance measure")) +
  theme(axis.text.x = element_text (angle = 60, hjust = 1))
```

## Performance measures of the JRip rules induced from 1,000 samples with 100 configurations

```{r JRIP-100-samples, warning=FALSE}
rule_results <- check_rule_induction_algorithms(
  sample_sizes=100,
  repetitions=1000,
  rule_induction_algorithms = "JRip",
)
desc <- describe(rule_results)
desc <- desc[-c(1, 2),]
desc <- desc[,-c(1, 2, 6, 7, 10, 11, 12, 13)]
print(desc, include.rownames=TRUE, digits=4)

tidied_rule_results <- rule_results %>%
  gather(Accuracy:Kappa, key = Measure, value = Value) %>%
  mutate(Measure = fct_recode(Measure,
                              "Recall/Sensitivity" = "Sensitivity"))
ggplot(tidied_rule_results, aes(x = Value)) +
  geom_density(alpha = 0.7, fill = "#3388CC", col="#3388CC") +
  facet_wrap( Measure ~ ., ncol=2) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05)) +
  labs(x = expression("Performance measure"),
       y = expression("Density")) +
  theme(axis.text.x = element_text (angle = 60, hjust = 1))

```
